---
title: "Yield monitor wrangling - 2020"
author: "Ujjwal Sigdel"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning objectives  
Today's learning objectives are:  

- Talk about the role of yield monitor data in ag  
- Import yield data  
- Explore with tables, plots, and maps (EDA)  
- Learn and implement data wrangling steps for geospatial data  

# Setup  
## Loading packages  
```{r setup }
#| message: false

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)

```

## Importing data  
```{r yield20 }
yield20 <- read_csv("../data/07 yield20.csv")

yield20 

```

# EDA  
```{r summary}
summary(yield20) 
```

What columns do you think are relevant?  

```{r yield distribution - density plot}
ggplot(yield20, aes(x=lb_ac))+
  geom_density()

```

```{r speed distribution - density plot}
ggplot(yield20, aes(x=Speed_mi_h))+
  geom_density()+
  geom_rug()

```

```{r yield map}
ggplot(yield20, aes(x= POINT_X,
                    y = POINT_Y))+
  geom_point(aes(color = lb_ac))+
  scale_color_viridis_b()

```

```{r yield restricted map}
yield20 %>% 
  filter(lb_ac < 3000) %>% 
ggplot(aes(x= POINT_X,
                    y = POINT_Y))+
  geom_point(aes(color = lb_ac))+
  scale_color_viridis_b()
  

```

```{r yield x speed}
ggplot(yield20, aes(x=Speed_mi_h, 
                    y=lb_ac))+
  geom_point()+
  geom_smooth()

```
Do we need to clean this data?  
From a yield map quality perspective, what steps would you do?  

- Clean based on speed  
- Clean based on yield  

```{r data needs}
yield20
```

From a data wrangling perspective, what steps would you do?  

- standardize column naming style (for ease of use)  
- select only needed columns:  

  - Speed  
  - Dry_Yield  
  - POINT_X  
  - POINT_Y  

- rename columns to include measurement units  
- create columns in metric system  
- transform it into spatial object (as of now, it is only a data frame)  
- transform from geographic (unprojected) to projected  

# Wrangling  
```{r yield20_w}
yield20_w <- yield20 %>% 
  # Cleaning names
  clean_names() %>% 
  # Selecting desired variables and renaming some of them
  dplyr::select(speed_mph = speed_mi_h,
                yield_lbac = lb_ac,
                point_x,
                point_y) %>% 
  # Transforming from english to metric units
  mutate(yield_kgha=yield_lbac*0.453592/0.404686,
         Speed_kmh=speed_mph*1.60934) %>% 

  # Transforming into sf object
  st_as_sf(coords = c("point_x", "point_y")) %>% 
  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>% 
  # Transforming from geodraphic to prjected CRS
  st_transform(crs = 6345) #32616?6345
yield20_w

```

# Maps  
```{r static}
yield20_w %>%
  filter(yield_lbac < 3000) %>%
  ggplot()+
  geom_sf(aes(color = yield_lbac), size = .1)+
  scale_color_viridis_b()

```

```{r interactive}
yield20_w %>%
  filter(yield_lbac < 3000) %>% 
  mapview(zcol = 'yield_lbac',
          alpha.regions = 1,
          cex = .1,
          lwd = 0,
          map.types = 'Esri.WorldImagery'
          )

```



# Exporting  
```{r}
write_sf(yield20_w, 
         "../data/yield20_processed.geojson",
         delete_dsn = T)

```


